---
title: "Codebook"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Source:
Jorge L. Reyes-Ortiz, Davide Anguita, Alessandro Ghio, Luca Oneto.
Smartlab - Non Linear Complex Systems Laboratory
DITEN - Università degli Studi di Genova.
Via Opera Pia 11A, I-16145, Genoa, Italy.
activityrecognition@smartlab.ws
www.smartlab.ws

## Data Set Information:
- The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) while wearing a smartphone (Samsung Galaxy S II) on the waist . Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 
- The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain. See 'features_info.txt' for more details. 

## Attribute Information:
For each record in the dataset it is provided:
- Triaxial acceleration from the accelerometer (total acceleration) and the estimated body acceleration.
- Triaxial Angular velocity from the gyroscope.
- A 561-feature vector with time and frequency domain variables.
- Its activity label.
- An identifier of the subject who carried out the experiment. 

## Feature Selection 
The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 
Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). These signals were used to estimate variables of the feature vector for each pattern:  
>'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

- tBodyAcc-XYZ
- tGravityAcc-XYZ
- tBodyAccJerk-XYZ
- tBodyGyro-XYZ
- tBodyGyroJerk-XYZ
- tBodyAccMag
- tGravityAccMag
- tBodyAccJerkMag
- tBodyGyroMag
- tBodyGyroJerkMag
- fBodyAcc-XYZ
- fBodyAccJerk-XYZ
- fBodyGyro-XYZ
- fBodyAccMag
- fBodyAccJerkMag
- fBodyGyroMag
- fBodyGyroJerkMag

The set of variables that were estimated from these signals are: 

- mean(): Mean value
- std(): Standard deviation
- mad(): Median absolute deviation 
- max(): Largest value in array
- min(): Smallest value in array
- sma(): Signal magnitude area
- energy(): Energy measure. Sum of the squares divided by the number of values. 
- iqr(): Interquartile range 
- entropy(): Signal entropy
- arCoeff(): Autorregresion coefficients with Burg order equal to 4
- correlation(): correlation coefficient between two signals
- maxInds(): index of the frequency component with largest magnitude
- meanFreq(): Weighted average of the frequency components to obtain a mean frequency
- skewness(): skewness of the frequency domain signal 
- kurtosis(): kurtosis of the frequency domain signal 
- bandsEnergy(): Energy of a frequency interval within the 64 bins of the FFT of each window.
- angle(): Angle between to vectors.

Additional vectors obtained by averaging the signals in a signal window sample. These are used on the angle() variable:

- gravityMean
- tBodyAccMean
- tBodyAccJerkMean
- tBodyGyroMean
- tBodyGyroJerkMean

```{r}
  ##Setup
  library("dplyr")
  
  ##Create Codebook sample
  f <- "C:\\Users\\Chris\\Documents\\R_Projects\\CourseProject\\tidyDataset.csv"
  data <- read.csv(f,header = TRUE, sep = ",")
  codeBook <- t(summary(data[4:ncol(data)]))
  
  write.table(codeBook,file = "codebook.csv", sep = ",")
  print(codeBook)
```

"PersonID"|"N/A"|"N/A"|"N/A"|"N/A"|"N/A"|"N/A"
Activities|N/A|N/A|N/A|N/A|N/A|N/A
tBodyAcc.mean...X|-1.0000  | 0.2626  | 0.2772  | 0.2743  | 0.2884  | 1.0000  
tBodyAcc.mean...Y|-1.00000  |-0.02490  |-0.01716  |-0.01774  |-0.01062  | 1.00000  
tBodyAcc.mean...Z|-1.00000  |-0.12102  |-0.10860  |-0.10892  |-0.09759  | 1.00000  
tBodyAcc.std...X|-1.0000  |-0.9924  |-0.9430  |-0.6078  |-0.2503  | 1.0000  
tBodyAcc.std...Y|-1.00000  |-0.97699  |-0.83503  |-0.51019  |-0.05734  | 1.00000  
tBodyAcc.std...Z|-1.0000  |-0.9791  |-0.8508  |-0.6131  |-0.2787  | 1.0000  
tGravityAcc.mean...X|-1.0000  | 0.8117  | 0.9218  | 0.6692  | 0.9547  | 1.0000  
tGravityAcc.mean...Y|-1.000000  |-0.242943  |-0.143551  | 0.004039  | 0.118905  | 1.000000  
tGravityAcc.mean...Z|-1.00000  |-0.11671  | 0.03680  | 0.09215  | 0.21621  | 1.00000  
tGravityAcc.std...X|-1.0000  |-0.9949  |-0.9819  |-0.9652  |-0.9615  | 1.0000  
tGravityAcc.std...Y|-1.0000  |-0.9913  |-0.9759  |-0.9544  |-0.9464  | 1.0000  
tGravityAcc.std...Z|-1.0000  |-0.9866  |-0.9665  |-0.9389  |-0.9296  | 1.0000  
tBodyAccJerk.mean...X|-1.00000  | 0.06298  | 0.07597  | 0.07894  | 0.09131  | 1.00000  
tBodyAccJerk.mean...Y|-1.000000  |-0.018555  | 0.010753  | 0.007948  | 0.033538  | 1.000000  
tBodyAccJerk.mean...Z|-1.000000  |-0.031552  |-0.001159  |-0.004675  | 0.024578  | 1.000000  
tBodyAccJerk.std...X|-1.0000  |-0.9913  |-0.9513  |-0.6398  |-0.2912  | 1.0000  
tBodyAccJerk.std...Y|-1.0000  |-0.9850  |-0.9250  |-0.6080  |-0.2218  | 1.0000  
tBodyAccJerk.std...Z|-1.0000  |-0.9892  |-0.9543  |-0.7628  |-0.5485  | 1.0000  
tBodyGyro.mean...X|-1.00000  |-0.04579  |-0.02776  |-0.03098  |-0.01058  | 1.00000  
tBodyGyro.mean...Y|-1.00000  |-0.10399  |-0.07477  |-0.07472  |-0.05110  | 1.00000  
tBodyGyro.mean...Z|-1.00000  | 0.06485  | 0.08626  | 0.08836  | 0.11044  | 1.00000  
tBodyGyro.std...X|-1.0000  |-0.9872  |-0.9016  |-0.7212  |-0.4822  | 1.0000  
tBodyGyro.std...Y|-1.0000  |-0.9819  |-0.9106  |-0.6827  |-0.4461  | 1.0000  
tBodyGyro.std...Z|-1.0000  |-0.9850  |-0.8819  |-0.6537  |-0.3379  | 1.0000  
tBodyGyroJerk.mean...X|-1.00000  |-0.11723  |-0.09824  |-0.09671  |-0.07930  | 1.00000  
tBodyGyroJerk.mean...Y|-1.00000  |-0.05868  |-0.04056  |-0.04232  |-0.02521  | 1.00000  
tBodyGyroJerk.mean...Z|-1.00000  |-0.07936  |-0.05455  |-0.05483  |-0.03168  | 1.00000  
tBodyGyroJerk.std...X|-1.0000  |-0.9907  |-0.9348  |-0.7313  |-0.4865  | 1.0000  
tBodyGyroJerk.std...Y|-1.0000  |-0.9922  |-0.9548  |-0.7861  |-0.6268  | 1.0000  
tBodyGyroJerk.std...Z|-1.0000  |-0.9926  |-0.9503  |-0.7399  |-0.5097  | 1.0000  
tBodyAccMag.mean..|-1.0000  |-0.9819  |-0.8746  |-0.5482  |-0.1201  | 1.0000  
tBodyAccMag.std..|-1.0000  |-0.9822  |-0.8437  |-0.5912  |-0.2423  | 1.0000  
tGravityAccMag.mean..|-1.0000  |-0.9819  |-0.8746  |-0.5482  |-0.1201  | 1.0000  
tGravityAccMag.std..|-1.0000  |-0.9822  |-0.8437  |-0.5912  |-0.2423  | 1.0000  
tBodyAccJerkMag.mean..|-1.0000  |-0.9896  |-0.9481  |-0.6494  |-0.2956  | 1.0000  
tBodyAccJerkMag.std..|-1.0000  |-0.9907  |-0.9288  |-0.6278  |-0.2733  | 1.0000  
tBodyGyroMag.mean..|-1.0000  |-0.9781  |-0.8223  |-0.6052  |-0.2454  | 1.0000  
tBodyGyroMag.std..|-1.0000  |-0.9775  |-0.8259  |-0.6625  |-0.3940  | 1.0000  
tBodyGyroJerkMag.mean..|-1.0000  |-0.9923  |-0.9559  |-0.7621  |-0.5499  | 1.0000  
tBodyGyroJerkMag.std..|-1.0000  |-0.9922  |-0.9403  |-0.7780  |-0.6093  | 1.0000  
fBodyAcc.mean...X|-1.0000  |-0.9913  |-0.9456  |-0.6228  |-0.2646  | 1.0000  
fBodyAcc.mean...Y|-1.0000  |-0.9792  |-0.8643  |-0.5375  |-0.1032  | 1.0000  
fBodyAcc.mean...Z|-1.0000  |-0.9832  |-0.8954  |-0.6650  |-0.3662  | 1.0000  
fBodyAcc.std...X|-1.0000  |-0.9929  |-0.9416  |-0.6034  |-0.2493  | 1.0000  
fBodyAcc.std...Y|-1.00000  |-0.97689  |-0.83261  |-0.52842  |-0.09216  | 1.00000  
fBodyAcc.std...Z|-1.0000  |-0.9780  |-0.8398  |-0.6179  |-0.3023  | 1.0000  
fBodyAcc.meanFreq...X|-1.00000  |-0.41878  |-0.23825  |-0.22147  |-0.02043  | 1.00000  
fBodyAcc.meanFreq...Y|-1.000000  |-0.144772  | 0.004666  | 0.015401  | 0.176603  | 1.000000  
fBodyAcc.meanFreq...Z|-1.00000  |-0.13845  | 0.06084  | 0.04731  | 0.24922  | 1.00000  
fBodyAccJerk.mean...X|-1.0000  |-0.9912  |-0.9516  |-0.6567  |-0.3270  | 1.0000  
fBodyAccJerk.mean...Y|-1.0000  |-0.9848  |-0.9257  |-0.6290  |-0.2638  | 1.0000  
fBodyAccJerk.mean...Z|-1.0000  |-0.9873  |-0.9475  |-0.7436  |-0.5133  | 1.0000  
fBodyAccJerk.std...X|-1.0000  |-0.9920  |-0.9562  |-0.6550  |-0.3203  | 1.0000  
fBodyAccJerk.std...Y|-1.0000  |-0.9865  |-0.9280  |-0.6122  |-0.2361  | 1.0000  
fBodyAccJerk.std...Z|-1.0000  |-0.9895  |-0.9590  |-0.7809  |-0.5903  | 1.0000  
fBodyAccJerk.meanFreq...X|-1.00000  |-0.29770  |-0.04544  |-0.04771  | 0.20447  | 1.00000  
fBodyAccJerk.meanFreq...Y|-1.000000  |-0.427951  |-0.236530  |-0.213393  | 0.008651  | 1.000000  
fBodyAccJerk.meanFreq...Z|-1.00000  |-0.33139  |-0.10246  |-0.12383  | 0.09124  | 1.00000  
fBodyGyro.mean...X|-1.0000  |-0.9853  |-0.8917  |-0.6721  |-0.3837  | 1.0000  
fBodyGyro.mean...Y|-1.0000  |-0.9847  |-0.9197  |-0.7062  |-0.4735  | 1.0000  
fBodyGyro.mean...Z|-1.0000  |-0.9851  |-0.8877  |-0.6442  |-0.3225  | 1.0000  
fBodyGyro.std...X|-1.0000  |-0.9881  |-0.9053  |-0.7386  |-0.5225  | 1.0000  
fBodyGyro.std...Y|-1.0000  |-0.9808  |-0.9061  |-0.6742  |-0.4385  | 1.0000  
fBodyGyro.std...Z|-1.0000  |-0.9862  |-0.8915  |-0.6904  |-0.4168  | 1.0000  
fBodyGyro.meanFreq...X|-1.00000  |-0.27189  |-0.09868  |-0.10104  | 0.06810  | 1.00000  
fBodyGyro.meanFreq...Y|-1.00000  |-0.36257  |-0.17298  |-0.17428  | 0.01366  | 1.00000  
fBodyGyro.meanFreq...Z|-1.00000  |-0.23240  |-0.05369  |-0.05139  | 0.12251  | 1.00000  
fBodyAccMag.mean..|-1.0000  |-0.9847  |-0.8755  |-0.5860  |-0.2173  | 1.0000  
fBodyAccMag.std..|-1.0000  |-0.9829  |-0.8547  |-0.6595  |-0.3823  | 1.0000  
fBodyAccMag.meanFreq..|-1.00000  |-0.09663  | 0.07026  | 0.07688  | 0.24495  | 1.00000  
fBodyBodyAccJerkMag.mean..|-1.0000  |-0.9898  |-0.9290  |-0.6208  |-0.2600  | 1.0000  
fBodyBodyAccJerkMag.std..|-1.0000  |-0.9907  |-0.9255  |-0.6401  |-0.3082  | 1.0000  
fBodyBodyAccJerkMag.meanFreq..|-1.000000  |-0.002959  | 0.164180  | 0.173220  | 0.357307  | 1.000000  
fBodyBodyGyroMag.mean..|-1.0000  |-0.9825  |-0.8756  |-0.6974  |-0.4514  | 1.0000  
fBodyBodyGyroMag.std..|-1.0000  |-0.9781  |-0.8275  |-0.7000  |-0.4713  | 1.0000  
fBodyBodyGyroMag.meanFreq..|-1.00000  |-0.23436  |-0.05210  |-0.04156  | 0.15158  | 1.00000  
fBodyBodyGyroJerkMag.mean..|-1.0000  |-0.9921  |-0.9453  |-0.7798  |-0.6122  | 1.0000  
fBodyBodyGyroJerkMag.std..|-1.0000  |-0.9926  |-0.9382  |-0.7922  |-0.6437  | 1.0000  
fBodyBodyGyroJerkMag.meanFreq..|-1.00000  |-0.01948  | 0.13625  | 0.12671  | 0.28896  | 1.00000  
angle.tBodyAccMean.gravity.|-1.000000  |-0.124694  | 0.008146  | 0.007705  | 0.149005  | 1.000000  
angle.tBodyAccJerkMean..gravityMean.|-1.000000  |-0.287031  | 0.007668  | 0.002648  | 0.291490  | 1.000000  
angle.tBodyGyroMean.gravityMean.|-1.00000  |-0.49311  | 0.01719  | 0.01768  | 0.53614  | 1.00000  
angle.tBodyGyroJerkMean.gravityMean.|-1.000000  |-0.389041  |-0.007186  |-0.009219  | 0.365996  | 1.000000  
angle.X.gravityMean.|-1.0000  |-0.8173  |-0.7156  |-0.4965  |-0.5215  | 1.0000  
angle.Y.gravityMean.|-1.000000  | 0.002151  | 0.182029  | 0.063255  | 0.250790  | 1.000000  
angle.Z.gravityMean.|-1.000000  |-0.131880  |-0.003882  |-0.054284  | 0.102970  | 1.000000  
